---
title: "Stratified analysis - Inj15"
format: docx
editor: visual
---

## 1. Install packages and load libraries

```{r}
# Load the required libraries into the current R session:
pacman::p_load(rio, 
               here, 
               tidyverse, 
               skimr,
               plyr,
               janitor,
               lubridate,
               gtsummary, 
               flextable,
               officer,
               epikit, 
               apyramid, 
               scales,
               EpiStats)
```

## 2. Import your data

```{r, Import_data}
# Import the raw data set: 
copdata <- rio::import(here::here("data", "Copenhagen_clean2.rds"))
```

## 3. Risk Ratio

### a) Veal as exposure of interest, stratified by having eaten pasta

```{r}
stratall <- copdata %>% 
  # Mutate across to convert cases to numeric:
  mutate(across(.cols = case, 
                .fns = ~ as.numeric(.)))

# Pass data to the csinter function:
pastastrata <- csinter(x = stratall, 
                       cases = "case", 
                       exposure = "veal", 
                       by = "pasta")

pastastrata
```

Let's check if pasta is associated with veal (if we are thinking veal may be a confounder, we need to see if there is an association between the potential confounder (veal) and the exposure (pasta)):

```{r}
# Perform Wilcoxon rank sum test on pasta and veal:
wilcox.test(pasta ~ veal, 
            data = copdata)
```

### c) Champagne as exposure of interest, stratified by having eaten pasta

```{r}
# Pass data to the csinter function:
champstrata <- csinter(x = stratall, 
                       cases = "case", 
                       exposure = "champagne", 
                       by = "pasta")
```
